On the uniqueness and stability of dictionaries for sparse representation of noisy signals

### 2017-04-07 manuscripts/PNAS/chazthm_pnas.tex
elaborated

\begin{corollary}\label{SLCopt}
If the assumptions of Thm.\ref{DeterministicUniquenessTheorem} hold with more than $(k-1)\left[ {\bar m \choose k} + |E|k{\bar m \choose k-1}\right]$ vectors $\mathbf{x}_i$ supported on each $S \in E$, then all solutions to Prob.~\ref{OptimizationProblem} necessarily satisfy implications 1) and 2) from Thm.~\ref{DeterministicUniquenessTheorem}.
\end{corollary}
%\begin{corollary}\label{SLCopt}
%If the assumptions of Thm.\ref{DeterministicUniquenessTheorem} hold with more than $(k-1)\left[ {\bar m \choose k} + |E|k{\bar m \choose k-1}\right]$ vectors $\mathbf{x}_i$ supported on each $S \in E$, then all matrices $\mathbf{B} \in \mathbb{R}^{n \times \bar m}$ and $\mathbf{\bar x}_1, \ldots, \mathbf{\bar x}_N \in \mathbb{R}^{\bar m}$ solving:
%\begin{align}\label{minsum}
%\min \sum_{i = 1}^N \|\mathbf{\bar x}_i\|_0 \ \ 
%\text{subject to} \ \ \|\mathbf{A}\mathbf{x}_i - \mathbf{B}\mathbf{\bar x}_i\|_2 \leq \varepsilon, \ \text{for all $i$},
%\end{align}
%
%necessarily satisfy implications 1) and 2) from Thm.~\ref{DeterministicUniquenessTheorem}.

@@

Since $f$ is a real analytic function, having one substitution of real numbers with $f(\mathbf{A}) \neq 0$ implies that its zeroes form a set of (Borel) measure zero. Fortunately, it is easy to construct such a matrix $\mathbf{A}$ by inserting $n-k$ rows of zeros into a $k \times m$ ``Vandermonde" matrix construction. %Let $\gamma_1, \ldots, \gamma_m$ be any distinct numbers. Then the columns of the $k \times m$ matrix $V = (\gamma^{\ell}_i)^{k,m}_{\ell,i=1}$ are in general linear position (since the $\gamma_i$ are distinct, any $k \times k$ ``Vandermonde" sub-determinant is nonzero). Concatenating $n-k$ rows of zeros to the bottom of this matrix yields a $n \times m$ matrix $\mathbf{A}$ satisfying the spark condition, i.e. with $f(\mathbf{A}) = 0$.
Hence, almost every real matrix $\mathbf{A}$ with at least $k$ rows and columns satisfies \eqref{SparkCondition}. %\eqref{CScondition} holds with a $\gamma$ for some distribution.  We set $\gamma_0$ to be the smallest known such $\gamma$. % We remark that the precise relationship between $m$, $n$, and $k$ guaranteeing that $f$ is not identically zero is a challenging problem in real algebraic geometry. In any case, we set $\gamma_0$ to be the smallest known such $\gamma$.

@@

\begin{corollary}\label{ProbabilisticCor}
Fix $n, m$ and $k$ such that $n \geq k < m$ % satisfying \eqref{CScondition} for $\gamma = \gamma_0$, 
and let the entries of $\mathbf{A} \in \mathbb{R}^{n \times m}$ and $k$-sparse $\mathbf{x}_1, \ldots, \mathbf{x}_N \in \mathbb{R}^m$ be drawn independently from probability measures absolutely continuous with respect to the standard Borel measure $\mu$. If more than $(k-1){m \choose k}$ of the vectors $\mathbf{x}_i$ are supported on each $S \in E$ for a regular $E \subseteq {[m] \choose k}$ with the SIP, then $Y$ has a stable $k$-sparse representation in $m$ dimensions with probability one.
\end{corollary}

@@

 \begin{proof}[Proof of Cor.~\ref{SLCopt}]
We bound the number of $k$-sparse $\mathbf{\bar x}_i$ and then apply Thm.~\ref{DeterministicUniquenessCorollary}. 
Let $n_p$ be the number of $\mathbf{\bar x}_i$ with $\|\mathbf{\bar x}_i\|_0 = p$.  %, so that $\sum_{p = 1}^{m} n_p = N$. 
Since the $\mathbf{x}_i$ are all $k$-sparse, by \eqref{minsum} we have:
%\begin{align}
\mbox{$k \sum_{p = 0}^{m} n_p \geq \sum_{i=0}^N \|\mathbf{x}_i\|_0 \geq \sum_{i=0}^N \|\mathbf{\bar x}_i\|_0 = \sum_{p=0}^{m} p n_p.$}
%\end{align}
Hence,
\begin{align}\label{eqn}
\sum_{p = k+1}^m n_p \leq \sum_{p = k+1}^m (p-k) n_p \leq \sum_{p = 0}^k (k-p)n_p \leq k \sum_{p = 0}^{k-1} n_p.
\end{align}

@@

Hence, the number of $(k-1)$-sparse vectors $\mathbf{\bar x}_i$ is no greater than $|E|(k-1){ m \choose k-1}$. Taking \eqref{eqn} into account, it follows that at least $N - |E|k(k-1){ m \choose k-1}$ of the vectors $\mathbf{\bar  x}_i$ are $k$-sparse. Therefore, more than $(k-1){m \choose k}$ of the vectors $\mathbf{x}_i$ supported on each $S \in E$ have corresponding $\mathbf{\bar x}_i$ that are also $k$-sparse. The result now follows directly from Thm.~\ref{DeterministicUniquenessCorollary}.
\end{proof}

>>>>

\begin{corollary}\label{SLCopt}
If the assumptions of Thm.\ref{DeterministicUniquenessTheorem} hold with more than $(k-1)\left[ {\bar m \choose k} + |E|k{m \choose k-1}\right]$ vectors $\mathbf{x}_i$ supported on each $S \in E$, then all solutions to Prob.~\ref{OptimizationProblem} necessarily satisfy implications 1) and 2) from Thm.~\ref{DeterministicUniquenessTheorem}.
\end{corollary}
%\begin{corollary}\label{SLCopt}
%If the assumptions of Thm.\ref{DeterministicUniquenessTheorem} hold with more than $(k-1)\left[ {\bar m \choose k} + |E|k{\bar m \choose k-1}\right]$ vectors $\mathbf{x}_i$ supported on each $S \in E$, then all matrices $\mathbf{B} \in \mathbb{R}^{n \times \bar m}$ and $\mathbf{\bar x}_1, \ldots, \mathbf{\bar x}_N \in \mathbb{R}^{\bar m}$ solving:
%\begin{align}\label{minsum}
%\min \sum_{i = 1}^N \|\mathbf{\bar x}_i\|_0 \ \ 
%\text{subject to} \ \ \|\mathbf{A}\mathbf{x}_i - \mathbf{B}\mathbf{\bar x}_i\|_2 \leq \varepsilon, \ \text{for all $i$},
%\end{align}
%
%necessarily satisfy implications 1) and 2) from Thm.~\ref{DeterministicUniquenessTheorem}.

@@

Since $f$ is a real analytic function, having one substitution of real numbers with $f(\mathbf{A}) \neq 0$ implies that its zeroes form a set of (Borel) measure zero. Fortunately, it is easy to construct such an $n \times m$ matrix $\mathbf{A}$ for any $n \geq \max(2k,m)$ by inserting enough rows of zeros into a $\max(2k, m) \times m$ ``Vandermonde" matrix construction. %Let $\gamma_1, \ldots, \gamma_m$ be any distinct numbers. Then the columns of the $k \times m$ matrix $V = (\gamma^{\ell}_i)^{k,m}_{\ell,i=1}$ are in general linear position (since the $\gamma_i$ are distinct, any $k \times k$ ``Vandermonde" sub-determinant is nonzero). Concatenating $n-k$ rows of zeros to the bottom of this matrix yields a $n \times m$ matrix $\mathbf{A}$ satisfying the spark condition, i.e. with $f(\mathbf{A}) = 0$.
Hence, almost every real $n \times m$ matrix $\mathbf{A}$ with $n \geq \max(2k,m)$ satisfies \eqref{SparkCondition}. %\eqref{CScondition} holds with a $\gamma$ for some distribution.  We set $\gamma_0$ to be the smallest known such $\gamma$. % We remark that the precise relationship between $m$, $n$, and $k$ guaranteeing that $f$ is not identically zero is a challenging problem in real algebraic geometry. In any case, we set $\gamma_0$ to be the smallest known such $\gamma$.

@@

\begin{corollary}\label{ProbabilisticCor}
Fix $k < m$ and $n \geq \max(2k, m)$ % satisfying \eqref{CScondition} for $\gamma = \gamma_0$, 
and let the entries of $\mathbf{A} \in \mathbb{R}^{n \times m}$ and $k$-sparse $\mathbf{x}_1, \ldots, \mathbf{x}_N \in \mathbb{R}^m$ be drawn independently from probability measures absolutely continuous with respect to the standard Borel measure $\mu$. If more than $(k-1){m \choose k}$ of the vectors $\mathbf{x}_i$ are supported on each $S \in E$ for a regular $E \subseteq {[m] \choose k}$ with the SIP, then $Y$ has a stable $k$-sparse representation in $m$ dimensions with probability one.
\end{corollary}

@@

\begin{proof}[Proof of Cor.~\ref{SLCopt}]
We bound the number of $k$-sparse $\mathbf{\bar x}_i$ and then apply Thm.~\ref{DeterministicUniquenessCorollary}. 
Let $n_p$ be the number of $\mathbf{\bar x}_i$ with $\|\mathbf{\bar x}_i\|_0 = p$.  %, so that $\sum_{p = 1}^{m} n_p = N$. 
Since the $\mathbf{x}_i$ are all $k$-sparse, by \eqref{minsum} we have:
%\begin{align}
\mbox{$k \sum_{p = 0}^{\bar m} n_p \geq \sum_{i=0}^N \|\mathbf{x}_i\|_0 \geq \sum_{i=0}^N \|\mathbf{\bar x}_i\|_0 = \sum_{p=0}^{\bar m} p n_p.$}
%\end{align}
Hence,
\begin{align}\label{eqn}
\sum_{p = k+1}^{\bar m} n_p \leq \sum_{p = k+1}^{\bar m} (p-k) n_p \leq \sum_{p = 0}^k (k-p)n_p \leq k \sum_{p = 0}^{k-1} n_p.
\end{align}

@@

Hence, the total number of $(k-1)$-sparse vectors $\mathbf{\bar x}_i$ is no greater than $|E|(k-1){ m \choose k-1}$. Taking \eqref{eqn} into account, no more than $|E|k(k-1){ m \choose k-1}$ vectors $\mathbf{\bar x}_i$ are \emph{not} $k$-sparse. Since for every $S \in E$ there are over $(k-1)\left[ {\bar m \choose k} + |E|k{ m \choose k-1} \right]$ vectors $\mathbf{x}_i$ supported there, it follows that more than $(k-1){\bar m \choose k}$ of them must have corresponding $\mathbf{\bar x}_i$ that are also $k$-sparse. The result now follows directly from Thm.~\ref{DeterministicUniquenessCorollary}.
\end{proof}
